Output Summary

### 2023-09-08 03-post-process/ms-content/ms-content.Rmd
ms-content: include a no-fall table
Co-author comments suggested figure was confusing. Now table shows raw values, not expressed as ratios

no_fall_caption = "
  Ratios between the across-year mean (dark grey) and coefficient of variation (light grey) abundance of natural-origin smolt reaching Lower Granite Dam with (denominator) and without (numerator) the fall migrant life history strategy.
  The 'without' scenario was calculated by assuming all parr recruits experienced freshwater survival rates of spring migrants.
  Mean ratios <1 and CV ratios >1 indicate LGR smolt abundance would be lower and more variable, respectively, in the absence of the fall migrant strategy.
  Values presented are posterior means and 95% credible regions.
"

@@

```{r no-fall-fig, fig.width = 3.4, fig.height = 3.4, fig.cap = msdown::prep_caption(no_fall_caption), time_it = TRUE, eval = params$no_fall_fig}
cv = function(x) {sd(x, na.rm = TRUE)/mean(x, na.rm = TRUE)}

@@

  # calculate the ratios between the actual and hypothetical
  mean_p_ratio = colMeans(Ma_tot_hyp_i/Ma_tot_real_i, na.rm = TRUE)
  cv_ratio = apply(Ma_tot_hyp_i, 2, cv)/apply(Ma_tot_real_i, 2, cv)

  names(mean_p_ratio) = paste0("mean_p_ratio[", 1:5, "]")
  names(cv_ratio) = paste0("cv_ratio[", 1:5, "]")
  c(mean_p_ratio, cv_ratio)
}

##### STEP 3: APPLY FUNCTION TO ALL POSTERIOR SAMPLES, FORMAT, AND SUMMARIZE #####

# apply
ratio_out = t(sapply(1:nrow(Pb), f))

# format
ratio_post = post_convert(cbind(postpack:::id_mat(post), ratio_out))

# summarize
mean_summ = post_summ(ratio_post, "mean")
cv_summ = post_summ(ratio_post, "cv")

@@

>>>>

no_fall_fig_caption = "
  Ratios between the across-year mean (dark grey) and coefficient of variation (light grey) abundance of natural-origin smolt reaching Lower Granite Dam with (denominator) and without (numerator) the fall migrant life history strategy.
  The 'without' scenario was calculated by assuming all parr recruits experienced freshwater survival rates of spring migrants.
  Mean ratios <1 and CV ratios >1 indicate LGR smolt abundance would be lower and more variable, respectively, in the absence of the fall migrant strategy.
  Values presented are posterior means and 95% credible regions.
"

no_fall_table_caption = "
  Inter-annual mean and coefficient of variation (\\%CV) of natural-origin smolt reaching Lower Granite Dam in the presence and (hypothetical) absence of the fall migrant life history type.
  Findings indicate that smolt abundance would be lower on average and more variable across years if the fall migrant strategy did not exist.
"

@@

```{r no-fall-analysis, time_it = TRUE, eval = params$no_fall_fig | params$no_fall_tab}
cv = function(x) {sd(x, na.rm = TRUE)/mean(x, na.rm = TRUE)}

@@

  # calculate the ratios between the actual and hypothetical
  mean_p_ratio_i = colMeans(Ma_tot_hyp_i/Ma_tot_real_i, na.rm = TRUE)
  cv_ratio_i = apply(Ma_tot_hyp_i, 2, cv)/apply(Ma_tot_real_i, 2, cv)

  # calculate interannual cv by population and scenario
  Ma_cv_hyp_i = apply(Ma_tot_hyp_i, 2, cv)
  Ma_cv_real_i = apply(Ma_tot_real_i, 2, cv)

  # calculate interannual mean by population and scenario
  Ma_mean_hyp_i = colMeans(Ma_tot_hyp_i, na.rm = TRUE)
  Ma_mean_real_i = colMeans(Ma_tot_real_i, na.rm = TRUE)

  # calculate percent changes
  Ma_mean_change_i = (Ma_mean_hyp_i - Ma_mean_real_i)/Ma_mean_real_i
  Ma_cv_change_i = (Ma_cv_hyp_i - Ma_cv_real_i)/Ma_cv_real_i

  # add names
  names(Ma_mean_hyp_i) = paste0("Ma_mean_hyp[", 1:5, "]")
  names(Ma_mean_real_i) = paste0("Ma_mean_real[", 1:5, "]")
  names(Ma_cv_hyp_i) = paste0("Ma_cv_hyp[", 1:5, "]")
  names(Ma_cv_real_i) = paste0("Ma_cv_real[", 1:5, "]")
  names(Ma_mean_change_i) = paste0("Ma_mean_change[", 1:5, "]")
  names(Ma_cv_change_i) = paste0("Ma_cv_change[", 1:5, "]")

  names(mean_p_ratio_i) = paste0("mean_p_ratio[", 1:5, "]")
  names(cv_ratio_i) = paste0("cv_ratio[", 1:5, "]")
  c(mean_p_ratio_i, cv_ratio_i, Ma_mean_hyp_i, Ma_mean_real_i, Ma_cv_hyp_i, Ma_cv_real_i, Ma_mean_change_i, Ma_cv_change_i)
}

# apply to each posterior sample
out = t(sapply(1:nrow(Pb), f))
out = post_convert(cbind(postpack:::id_mat(post), out))

```{r no-fall-fig, fig.width = 3.4, fig.height = 3.4, fig.cap = msdown::prep_caption(no_fall_fig_caption), eval = params$no_fall_fig}
# summarize
mean_summ = post_summ(out, "mean_p_ratio")
cv_summ = post_summ(out, "cv_ratio")

@@

`r ifelse(params$no_fall_fig & params$no_fall_tab, "_**Alternatively**, as a table, showing the actual numbers and a percent change, rather than as ratios._", "")`

```{r no-fall-tab, eval = params$no_fall_tab}
f = function(x, is_percent) {
  g = function(x, is_percent) {
    if (!is_percent) {
      prettyNum(round(x, 1), big.mark = ",", scientific = FALSE)
    } else {
      paste0(round(x, 2) * 100, "\\%")
    }
  }
  paste0(g(x["mean",], is_percent), " (", g(x["sd",], is_percent), ")")
}

tab = cbind(
  pop = c("CAT", "LOS", "MIN", "UGR", "Total"),
  f(post_summ(out, "Ma_mean_real")/1000, FALSE),
  f(post_summ(out, "Ma_mean_hyp")/1000, FALSE),
  f(post_summ(out, "Ma_mean_change"), TRUE),
  f(post_summ(out, "Ma_cv_real"), TRUE),
  f(post_summ(out, "Ma_cv_hyp"), TRUE)
)

kbl(tab, "latex", booktabs = TRUE, 
    col.names = c(
      " ", "Real\\textsuperscript{1,3}", "All Spring\\textsuperscript{2,3}", "\\%$\\Delta$",
      "Real\\textsuperscript{1}", "All Spring\\textsuperscript{2}"), escape = FALSE,
    caption = msdown::prep_caption(no_fall_table_caption)) |>
  kable_styling(full_width = FALSE, latex_options = "HOLD") |>
  add_header_above(c(" " = 1, "Inter-Annual Mean" = 3, "Inter-Annual \\\\%CV" = 2), bold = TRUE, escape = FALSE) |>
  column_spec(1, bold = TRUE) |>
  row_spec(0, bold = TRUE) |>
  footnote(general_title = "",
           general = c(
             "\\\\textit{Posterior summaries represented as mean (standard deviation).}"
           ),
           number =  c(
             "Based on the observed time series, in which both spring and fall migrants were present.",
             "Based on the hypothetical case where no parr recruits were fall migrants -- calculated after model fitting by assuming all parr recruits experience the higher mortality rates of the spring migrant life history type.",
             "Abundance expressed as thousands of natural-origin smolt reaching Lower Granite Dam."
           ), threeparttable = TRUE, escape = FALSE)
